Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-02 paper-2024-sac-marl-neighbouring-strategies.tex
chore: add a todo on centralized actor critic approaches

Several algorithms have attempted to address these coordination challenges within the independent learner paradigm. 
Basic decentralized Q-learning~\cite{tan1993multi} allows each agent to learn its Q-values independently but often converges to suboptimal joint policies due to a lack of coordination. 
Subsequent work introduced heuristic-based extensions to improve coordination.
Approaches like distributed Q-learning~\cite{lauer2000algorithm} incorporate optimism in Q-value updates and equilibrium selection mechanisms to mitigate the convergence to shadowed equilibria (i.e., equilibria that are unstable due to the independent learning processes). 
Instead, algorithms such as hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient} and win-or-learn fast policy hill climbing~\cite{bowling2002multiagent} employ adaptive learning rates to address the issue of alternating exploration between agents, improving robustness and convergence.
%
Empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
While initially developed for tabular settings, some of these methods, like hysteretic Q-learning~\cite{palmer2017lenient} and lenient Q-learning, have been extended to deep reinforcement learning settings. 
However, these independent approaches, even with deep learning, often exhibit instability and slow convergence in complex environments and fundamentally lack explicit mechanisms for inter-agent coordination. 
Consequently, more sophisticated MARL algorithms have been developed that explicitly account for the interactions and policies of other agents to facilitate more effective coordination.
\paragraph{Neighboring-Based Methods}
\paragraph{Centralized Actor-Critic Methods}

\subsection{Motivation}
MADRL faces a critical challenge in balancing scalability and performance.  
Centralized training methods, while effective, often suffer from computational bottlenecks and difficulties scaling to large numbers of agents. 
While distributed training approaches offer improved scalability, 
they frequently assume independent learners, 
neglecting crucial inter-agent coordination and leading to suboptimal global performance.  
%
Neighbor-based methods have shown promise in simpler, tabular settings by enabling localized coordination,
but their integration with deep reinforcement learning remains largely unexplored. 
This work bridges this gap by introducing a novel approach that combines the power of deep learning with the scalability of neighbor-based communication.  
Our method aims to achieve both efficient decentralized learning and effective coordination, enabling online adaptation and robust performance in complex multi-agent environments.
\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
This section introduces three neighboring-based distributed learning strategies for multi-agent reinforcement learning. 
These strategies leverage local interactions between agents to improve learning efficiency and coordination.
For this investigation, we focus on deep Q-learning as the underlying learning algorithm,
and we consider the following strategies:
 k-nearest neighbor averaging, k-nearest neighbor consensus, and experience sharing.

>>>>

Several algorithms have attempted to address these coordination challenges within the independent learner paradigm. 
Basic decentralized Q-learning~\cite{tan1993multi} allows each agent to learn its Q-values independently but often converges to suboptimal joint policies due to a lack of coordination. 
Subsequent work introduced heuristic-based extensions to improve coordination.
Approaches like distributed Q-learning~\cite{lauer2000algorithm} incorporate optimism in Q-value updates and equilibrium selection mechanisms to mitigate the convergence to shadowed equilibria (i.e., equilibria that are unstable due to the independent learning processes). 
Instead, algorithms such as hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient} and win-or-learn fast policy hill climbing~\cite{bowling2002multiagent} employ adaptive learning rates to address the issue of alternating exploration between agents, improving robustness and convergence.
%
Empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
While initially developed for tabular settings, some of these methods, like hysteretic Q-learning~\cite{palmer2017lenient} and lenient Q-learning, have been extended to deep reinforcement learning settings. 
However, these independent approaches, even with deep learning, often exhibit instability and slow convergence in complex environments and fundamentally lack explicit mechanisms for inter-agent coordination. 
Consequently, more sophisticated MARL algorithms have been developed that explicitly account for the interactions and policies of other agents to facilitate more effective coordination.
\paragraph{Neighboring-Based Methods}
Neighboring-based methods in MARL address the scalability and communication overhead challenges of fully centralized approaches by restricting agent interactions to a local neighborhood. 
This introduces a structure where agents communicate and coordinate with a limited set of neighbors, reducing the complexity of joint action and policy learning.
%
One prominent category within neighboring-based methods is consensus-based algorithms. 
These algorithms aim to reach a consensus among agents on shared values or policies through local information exchange. 
Early work like QD-learning~\cite{kar2012qd} extends Q-learning with a consensus update term, 
enabling agents to converge towards optimal Q-values despite only receiving local rewards.
%  
Recent advances in consensus-based MARL leverage deep learning and actor-critic architectures, 
combining local policy updates with consensus updates on critic parameters or gradients, 
as seen in the recent works~\cite{7434032,pennesi2010distributed,zhang2018fully,DBLP:journals/jmlr/CiosekW20}.

These algorithms often exhibit improved scalability and communication efficiency compared to fully centralized methods, 
but still need lack in capturing the global structure of the environment and the interactions between agents.
\paragraph{Centralized Actor-Critic Methods}
\gianlu{
  todo -- discuss the main three solution, policy based (MAPPO, MADDPG, etc.), value based (QMIX, VDQN, etc.) and mean field (mean-q learning with field-informed RL).
}
\subsection{Motivation}
MADRL faces a critical challenge in balancing scalability and performance.  
Centralized training methods, while effective, often suffer from computational bottlenecks and difficulties scaling to large numbers of agents. 
While distributed training approaches offer improved scalability, 
they frequently assume independent learners, 
neglecting crucial inter-agent coordination and leading to suboptimal global performance.  
%
Neighbor-based methods have shown promise in simpler, tabular settings by enabling localized coordination,
but their integration with deep reinforcement learning remains largely unexplored. 
This work bridges this gap by introducing a novel approach that combines the power of deep learning with the scalability of neighbor-based communication.  
Our method aims to achieve both efficient decentralized learning and effective coordination, enabling online adaptation and robust performance in complex multi-agent environments.
\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
This section introduces three neighboring-based distributed learning strategies for multi-agent reinforcement learning. 
These strategies leverage local interactions between agents to improve learning efficiency and coordination.
For this investigation, we focus on deep Q-learning as the underlying learning algorithm,
and we consider the following strategies:
 k-nearest neighbor averaging, k-nearest neighbor consensus, and experience sharing.
